---
title: "MA-MALDI-TOF: Machine-learning Algorithm based MALDI-TOF"
subtitle: 'Data analytics project'
author(s): "David Ayodele, Exauce Saya"
date: "12/12/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Overview

In lecture, we discussed regularization and the relationship between Bayesian linear models and Ridge regression. You will build upon the discussion from lecture by focusing on Lasso regression in this assignment.  

## Load packages

```{r, load_packages, eval=TRUE}
library(dplyr)
library(ggplot2)
library(plyr)
library(readr)
library(stringr)
library(gsubfn)
```

```{r}
file_list = list('Deinococcus.aquaticus_FR100+Run2-A^1.txt',
'Deinococcus.aquaticus_FR100+Run2-A^2.txt',
'Deinococcus.aquaticus_FR100+Run2-A^3.txt',
'Deinococcus.aquaticus_P17+Run2-A^10.txt',
'Deinococcus.aquaticus_P17+Run2-A^11.txt',
'Deinococcus.aquaticus_P17+Run2-A^12.txt',
'Deinococcus.aquaticus_P1+Run2-A^1.txt',
'Deinococcus.aquaticus_P1+Run2-A^2.txt',
'Deinococcus.aquaticus_P1+Run2-A^3.txt',
'Deinococcus.aquaticus_P21+Run2-B^1.txt',
'Deinococcus.aquaticus_P21+Run2-B^2.txt',
'Deinococcus.aquaticus_P21+Run2-B^3.txt',
'Deinococcus.aquaticus_P22+Run2-B^4.txt',
'Deinococcus.aquaticus_P22+Run2-B^5.txt',
'Deinococcus.aquaticus_P22+Run2-B^6.txt',
'Deinococcus.aquaticus_P23+Run2-B^7.txt',
'Deinococcus.aquaticus_P23+Run2-B^8.txt',
'Deinococcus.aquaticus_P23+Run2-B^9.txt',
'Deinococcus.aquaticus_P2+Run2-A^4.txt',
'Deinococcus.aquaticus_P2+Run2-A^5.txt',
'Deinococcus.aquaticus_P2+Run2-A^6.txt',
'Deinococcus.aquaticus_P34+Run2-B^10.txt',
'Deinococcus.aquaticus_P34+Run2-B^11.txt',
'Deinococcus.aquaticus_P34+Run2-B^12.txt',
'Deinococcus.aquaticus_P41+Run2-C^1.txt',
'Deinococcus.aquaticus_P41+Run2-C^2.txt',
'Deinococcus.aquaticus_P41+Run2-C^3.txt',
'Deinococcus.aquaticus_P43+Run2-C^7.txt',
'Deinococcus.aquaticus_P43+Run2-C^8.txt',
'Deinococcus.aquaticus_P43+Run2-C^9.txt',
'Deinococcus.aquaticus_P49+Run2-C^10.txt',
'Deinococcus.aquaticus_P49+Run2-C^11.txt',
'Deinococcus.aquaticus_P49+Run2-C^12.txt',
'Deinococcus.aquaticus_P65+Run2-A^1.txt',
'Deinococcus.aquaticus_P65+Run2-A^2.txt',
'Deinococcus.aquaticus_P65+Run2-A^3.txt',
'Deinococcus.aquaticus_P71+Run2-A^1.txt',
'Deinococcus.aquaticus_P71+Run2-A^2.txt',
'Deinococcus.aquaticus_P71+Run2-A^3.txt',
'Deinococcus.aquaticus_P72+Run2-A^1.txt',
'Deinococcus.aquaticus_P72+Run2-A^2.txt',
'Deinococcus.aquaticus_P72+Run2-A^3.txt',
'Deinococcus.aquaticus_P74+Run2-A^1.txt',
'Deinococcus.aquaticus_P74+Run2-A^2.txt',
'Deinococcus.aquaticus_P74+Run2-A^3.txt',
'Deinococcus.aquaticus_P76+Run2-A^1.txt',
'Deinococcus.aquaticus_P76+Run2-A^2.txt',
'Deinococcus.aquaticus_P76+Run2-A^3.txt',
'Deinococcus.aquaticus_P79+Run2-A^1.txt',
'Deinococcus.aquaticus_P79+Run2-A^2.txt',
'Deinococcus.aquaticus_P79+Run2-A^3.txt',
'Deinococcus.aquaticus_P7+Run2-A^7.txt',
'Deinococcus.aquaticus_P7+Run2-A^8.txt',
'Deinococcus.aquaticus_P7+Run2-A^9.txt',
'Deinococcus.aquaticus_P80+Run2-A^1.txt',
'Deinococcus.aquaticus_P80+Run2-A^2.txt',
'Deinococcus.aquaticus_P80+Run2-A^3.txt',
'Deinococcus.aquaticus_P81+Run2-A^1.txt',
'Deinococcus.aquaticus_P81+Run2-A^2.txt',
'Deinococcus.aquaticus_P81+Run2-A^3.txt',
'Deinococcus.gobiensis_Gobi+Run2-A^1.txt',
'Deinococcus.gobiensis_Gobi+Run2-A^2.txt',
'Deinococcus.gobiensis_Gobi+Run2-A^3.txt')
```



```{r}
file_list[10]
```
for (i in 1:length(file_list)) assign(temp[i], read.csv(temp[i]))
# dat_txt = ldply(myfiles, read.table, sep = "\t", fill=TRUE, header = TRUE)
```{r}
for (i in 1:length(file_list)) {
  data_3D[i] = readr::read_table2(str_c("https://raw.githubusercontent.com/davidayodele/ma-maldi-tof/master/RAWSPECTRA/",file_list[i]), " ")
}

#data_3D[2]
```

```{r}
#data_3D[[1]][2] # col 1 row 2
data_3D[[1]][[2]] # col 2 row 1
```


```{r}
dim(data_3D[[1]])
```


col.names <- c("group", "x1", "x2", "x3")
sep <- ","  # if comma can appear in fields use something else
pat <- "^(.*) +(\\d+) +(\\d+) +(\\d+) *$"

tmp <- sapply(strapplyc(r, pat), paste, collapse = sep)
read.table(text = tmp, col.names = col.names, as.is = TRUE, sep = sep)
```{r}

colnames <- c("Intensity", "m2z")
sep <- ","  # if comma can appear in fields use something else
pat <- "^(\\d+) +(\\d+) *$"


tmp <- sapply(strapplyc(data_3D[1], pat), paste, collapse = sep)
read.table(text = tmp, col.names = colnames, as.is = TRUE, sep = sep)

#colnames(data_3D[1]) <- c("Intensity","m2z")

```

```{r}
ggplot(mapping = aes(x = data_3D[1][])) +
  geom_line(size = 1.15, color = "black",
            mapping = aes(y = data_3D[1]))
```


```{r}
data1 <- readr::read_delim("https://raw.githubusercontent.com/davidayodele/ma-maldi-tof/master/RAWSPECTRA/Deinococcus.aquaticus_FR100%2BRun2-A%5E1.txt", " ")

data1 %>% glimpse()
```

## Creates a 4D array buffer with 1,000 rows, 1,000 cols, 10 sets, 10 labels (each set of 2D data is labeled)

```{r}
buffer_3D <- rep(0, 1000*1000*10); 

data_3D <- array(buffer_3D, c(1000, 1000, 10)); 
```

ar[304,5,2] <- 204;
```{r}
data_3D[,,1] = data1[1:1000,1:1000]

data_3D
```
```{r}
data_3D
```






```{r}
data2 <- readr::read_delim("https://raw.githubusercontent.com/davidayodele/ma-maldi-tof/master/RAWSPECTRA/Deinococcus.aquaticus_FR100%2BRun2-A%5E2.txt", " ")

data2 %>% glimpse()
```


```{r}
data_3D[,,2] = data2[1:1000, 1:1000]
```


